---
title: "Statistics"
output:
    html_document:
      toc: true
      toc_float: true

---

Some commonly used statistics calculations for use in RStudio.

# Odds ratio & Relative risk

Using epitools - [manual here](https://cran.r-project.org/web/packages/epitools/epitools.pdf)

*Reminder:* If you need more information on the tests use the help command in the console (e.g. `?riskratio`, `?oddsratio`).

**Library**
```{r odds-ratio1, message=FALSE, warning=FALSE}
library(epitools)
# if you don't have this package, first install using `install.packages("epitools")`
```

Create a simple dataframe. In this case we'll test effect of gender on parasite presence with a simple positive/negative summary. Of course if you have a your raw data in a spreadsheet you could make your own by summarising the releavnt information into a dataframe. (Need help tidying and summarising your data...check out [this tutorial](https://datacarpentry.org/R-ecology-lesson/03-dplyr.html) to check you hooked on the `dplyr` and `tidyr` packages 
```{r odds-ratio2, message=FALSE, warning=FALSE}
factor1 <- c("Female", "Male")
factor2 <- c("Positive", "Negative")
```

```{r odds-ratio3, message=FALSE, warning=FALSE}
dat <- matrix(c(16, 30, 15, 34), nrow = 2, ncol = 2, byrow = TRUE)
dimnames(dat) <- list("Sex" = factor1, "Parasite present" = factor2)
```

Your dataframe should look like this
```{r odds-ratio4}
dat
```

Now lets calculate our **odds ratio**
```{r odds-ratio5}
oddsratio(dat)
```

and **relative risk**
```{r odds-ratio6}
riskratio(dat)
```