---
title: "Mapping"
output:
  html_document:
    toc: true
    toc_float: true
---

# Maping option 1

Website
http://www.flutterbys.com.au/stats/tut/tut5.4.html

## Load libraries
```{r}
library(maps)
library(mapdata)
library(sp)
library(dplyr)
library(viridis)
library(readr)
library(ggplot2)
library(plotly)
```

<div class="alert alert-warning"
  <strong>Make sure the `tidyverse` package is not loaded. `ggplot2` and `dplyr` are okay.</strong>
</div>

## Map your points
Map your points (lat, lon) on a map of Australia

Load your point data file
```{r}
obs.data <- read_csv(file = "data/sites.csv")
```



Load Australian made, and refine to include Tasmania only
```{r}
tas<-map("worldHires", "Australia", fill=TRUE, xlim=c(144,150),
         ylim=c(-44,-40), mar=c(0,0,0,0))
```
Function to convert map data to SpatialPolygons
```{r}
map2SpatialPolygons <- function(df, proj4string=CRS("+proj=longlat")) {
  Plys <- list()
  i<-1
  mtch <- which(is.na(df$x))
  if(length(mtch)==0) {
    mtch <- length(df$x)+1
  } else {mtch <-mtch}
  shps <- length(mtch)
  #make sure the names are unique
  nms <- df$names
  nms[duplicated(nms)] <- paste(nms[duplicated(nms)],1:length(nms[duplicated(nms)]))
  for (j in 1:shps){
    Plys[[j]] <- Polygons(list(Polygon(cbind(df$x[i:(mtch[j]-1)],
                                             df$y[i:(mtch[j]-1)]))),ID=nms[j])
    i <- mtch[j]+1
  }
  SpatialPolygons(Plys,proj4string=proj4string)
}
tas.sp <- map2SpatialPolygons(tas)
par(mar=c(0,0,0,0))
plot(tas.sp, asp=1)
```
***


Plotting via ggplot
```{r}
g1 = ggplot(fortify(tas.sp), aes(y=lat, x=long, group=group), fill = palegreen) + geom_polygon()
g2 = ggplot(data = tas.sp) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill = NA, color = "black") + 
  coord_fixed(1.3) + theme_classic() 
```

```{r}
g3 <- g2 + geom_point(data = obs.data, aes (x=Lon, y=Lat, shape=Locality, stroke = 4)) + scale_colour_viridis_d(option = "inferno") + theme(
  axis.text.x = element_blank(),
  axis.text.y = element_blank(),
  axis.ticks = element_blank(),
  axis.line = element_blank(),
  axis.title = element_blank())
```

View plot
```{r}
g3
```

Save plot
```{r}
pdf("tas_map.pdf")
print(g3)
dev.off()
```

***

# Mapping option 2

Here is a second option for Mapping Australia.

In the above code the Australia map doesn't come with state lines. By downloading the mapping files below we can customise things a bit more.

Here, lets just make a map of Australia

Following tutorial at this [link](http://environmentalcomputing.net/making-maps-of-your-study-sites/)

```{r}
library(sp)
library(dplyr)
library(rgdal)
library(raster)
library(ggplot2)
library(readr)
```

xlim=c(110,160),
         ylim=c(-45,-5)

Create buffer around your data points
```{r}
buffer <- 0.11
min_lon <- 110
max_lon <- 160
min_lat <- -45
max_lat <- -5

geo_bounds <- c(left = min_lon, bottom = min_lat, right = max_lon, top = max_lat)
```

```{r}
Sites.grid <- expand.grid(lon_bound = c(geo_bounds[1], geo_bounds[3]), 
                       lat_bound = c(geo_bounds[2], geo_bounds[4]))
coordinates(Sites.grid) <- ~ lon_bound + lat_bound
```


**Load mapping files**

You will need to download the mapping files by following the link to website -[GEODATA COAST 100K 2004 [shapefiles](https://data.gov.au/dataset/geodata-coast-100k-2004/resource/455ee5c6-2862-445e-8fbe-604476926b4c)

```{r}
Aus <- readOGR(dsn = "~/Documents/Programs/R/Maps/Rpackage_Geodata-coast/australia",layer = "cstauscd_r")
```

Remove coastline clutter from map
```{r}
Aus_coast <- subset(Aus, FEAT_CODE != "sea")
plot <- plot(Aus_coast)
```

```{r}
Aus_crop <- crop(Aus_coast, extent(Sites.grid))
plot(Aus_crop)
```

plot

```{r}
aus_plot1 <- ggplot() + 
 geom_polygon(data = Aus_crop, aes(x=long, y=lat, group=group), fill = NA, colour="black") +
 coord_equal() +
 
 labs(x="Longitude", y="Latitude") +
 theme_classic() +
  theme_classic()
```

Save
```{r}
pdf("aus_map.pdf")
print(aus_plot1)
dev.off()
```

